### 2019-10-13 unpublished/notation clarification/notesOnNotation.tex

% Document settings
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{setspace}
\usepackage{amsmath,amssymb,gensymb,epsfig, epstopdf}
\usepackage{xcolor}
\pagestyle{plain}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}

\begin{document}

% Header
\textbf{Nitin Kapania} \\
\textbf{Notes on Vehicle Dynamics Notation} \\

The vehicle dynamics notes previously introduced use the following notation:

\begin{center}
\begin{tabular}{ l | r }
 $v_y$ & Vehicle lateral velocity in body-fixed frame \\ 
 $v_x$ & Vehicle longitudinal velocity in body-fixed frame \\ 
 $w_z$ & Vehicle yaw rate \\   
\end{tabular}
\end{center}

For the next two lectures, we will use the following notation:

\begin{center}
\begin{tabular}{ l | r }
 $U_y$ & Vehicle lateral velocity in body-fixed frame \\ 
 $U_x$ & Vehicle longitudinal velocity in body-fixed frame \\ 
 $r$ & Vehicle yaw rate \\   
\end{tabular}
\end{center}

\end{document}